---
title: "Measuring Economic Policy Uncertainty for the Israeli Economy"
author: "Yaakov Chen-Zion, Michael Gurkov and Nimrod Segev"
institute: "Bank of Israel \\newline Research Department \\newline \\newline Any views expressed in the Discussion Paper are those of the authors and do not necessarily reflect those of the Bank of Israel"
date: "13 December, 2021"
classoption: t
output:
  beamer_presentation:
    includes:
      in_header: !expr here::here('PresentationPreamble.tex')
    latex_engine: xelatex
    slide_level: 2
# bibliography: "GaR.bib"
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, echo = FALSE, message = FALSE,
  warning = FALSE, comment = "#>"

)
```


```{r load_libraries}


library(lubridate)

library(glue)

library(xts)

library(slider)

library(readxl)

library(corrr)

library(vars)

library(aTSA)

library(tidyverse)

library(slider)

```


```{r set_params}

events_df = tribble(
  ~ event,
  ~ category,
  ~ start_date,
  ~ end_date,
  "Intifada",
  "crisis",
  "2001-01-01",
  "2003-01-01",
  "GFC",
  "crisis",
  "2008-01-01",
  "2009-01-01",
  "Corona",
  "crisis",
  "2020-01-01",
  "2021-01-01",
  "Ariel Sharon",
  "political",
  "2006-01-01",
  "2006-02-01",
  "social justice protests",
  "political",
  "2011-08-01",
  "2011-12-01",
  "Brexit",
  "international",
  "2016-06-01",
  "2016-07-01",
  "Trump",
  "international",
  "2016-11-01",
  "2016-12-01"
) %>%
  mutate(across(ends_with("date"), as_date))


```


```{r import_data_epu}

epu_month = read_csv(glue(Sys.getenv("USERPROFILE"),
                       "\\OneDrive - Bank Of Israel\\Data\\",
                       "EPU\\TDM-JP\\temp_epu_month.csv")) %>% 
  dplyr::select(date = time_period, epu = epu_scaled)


epu_month_topics = read_csv(glue(Sys.getenv("USERPROFILE"),
                       "\\OneDrive - Bank Of Israel\\Data\\",
                       "EPU\\TDM-JP\\temp_epu_month_topics.csv")) %>% 
  dplyr::select(date = time_period, monetary_epu = monetary_epu_scaled,
         fiscal_epu = fiscal_epu_scaled) %>%
  mutate(date = as_date(paste0(date,"-01"))) %>% 
  pivot_longer(-date)



```

```{r import_epus}

url_table = tribble(
  ~ name, ~ file_name,~ var_name,
  "usa","US_Policy_Uncertainty_Data","News_Based_Policy_Uncert_Index",
  "global","Global_Policy_Uncertainty_Data", "GEPU_current",
  "euro","Europe_Policy_Uncertainty_Data","European_News_Index"
)

epus_df = pmap(url_table,function(name, file_name, var_name){
  
  temp_df = read_xlsx(glue(Sys.getenv("USERPROFILE"),
                           "\\OneDrive - Bank Of Israel\\Data\\EPU",
                           "\\epus\\",file_name, ".xlsx"))
  
  temp_df = temp_df %>% 
    mutate(Month = str_pad(Month,side = "left",width = 2,pad = "0")) %>% 
    mutate(date = as_date(paste(Year,Month,"01", sep = "-"))) %>% 
    dplyr::select(date, !!sym(name) := all_of(var_name))
      
  
  return(temp_df)

  
}) %>% 
  reduce(full_join, "date")

rm(url_table)

```

```{r import_vix}

vix = read_csv(glue(Sys.getenv("USERPROFILE"),
                    "\\OneDrive - Bank Of Israel\\Data",
                    "\\TASE\\vix_israel.csv")) %>% 
  set_names(c("date","vix")) %>% 
  mutate(date = dmy(date))

vix = vix %>% 
  group_by(date = as.yearmon(date)) %>%
  summarise(vix = mean(vix, na.rm = TRUE), .groups = "drop") %>%
  mutate(date = paste("01", str_pad(month(date),2,"left",0),
                           year(date),sep = "-")) %>%
  mutate(date = dmy(date))

```

```{r calculate_benchmark_df}

benchmark_df = epu_month %>%
  full_join(epus_df, "date") %>%
  full_join(vix, "date")


```


```{r import_var_data}

vars_list = c("epu", "i_il", "emp_il_sa",
              "gdp_il_sa","fx_eff")

var_df = read_excel(glue(Sys.getenv("USERPROFILE"),
                         "\\OneDrive - Bank Of Israel",
                         "\\Data\\EPU\\EPUI",
                         "\\EPUI_Data_activity_ALL_q.xlsx"),
                    sheet = "DATA")

var_df = var_df %>%
  mutate(date = as.yearqtr(paste0(year,"Q",quarter))) %>% 
  select(date,any_of(vars_list),starts_with("fx")) %>%
  mutate(fx_eff = if_else(year(date) > 1999,
                          fx_ner_ef,
                          0.75 * log(fx_us) + 0.07 * fx_japan + 0.18 * fx_uk)) %>%
  mutate(across(any_of(vars_list),log,.names = "{.col}_log")) %>%
  mutate(across(ends_with("_log"), ~ . - lag(.),.names = "{.col}_diff")) %>% 
  filter(date >= as.yearqtr("1995Q2") & year(date) <= 2019)


```

```{r run_var, cache=TRUE}

irf_df =  var_df  %>%
  dplyr::select(epu_log, fx_eff_log, i_il_log, emp_il_sa_log, gdp_il_sa_log)  %>%
  ts(start = c(1995, 2), frequency = 4)  %>%
  VAR(p = 8, ic = "AIC", type = "both", lag.max = 8)  %>%
  irf(n.ahead = 6,runs = 1000,ci = 0.90,impulse = "epu_log")

var_irf_delta =  var_df  %>%
  dplyr::select(paste0(vars_list,"_log_diff")) %>%
  dplyr::rename_all(~str_remove(.,"_log_diff")) %>% 
  filter(complete.cases(.)) %>% 
  ts(start = c(1995, 2), frequency = 4)  %>%
  VAR(p = 8, ic = "AIC", type = "both", lag.max = 8)  %>%
  irf(n.ahead = 6,cumulative = TRUE,runs = 1000,ci = 0.90,impulse = "epu")



```

## Overview

\vfill

\begin{itemize}
\setlength\itemsep{2em}
  \item
  Develop a granular EPU index for Israel following 
  Baker, Bloom, and Davis (2016)
  \begin{itemize}
    \item
    Monthly / Quarterly
    \item
    Subcategories
  \end{itemize}
  \item
  Convince you that the index makes sense
  \begin{itemize}
    \item
    responds to major local and international events
    \item
    correlates with alternative uncertainty measures
  \end{itemize}
  \item
  Examine the connection between the Israeli EPU and macroeconomic variables
  \begin{itemize}
    \item
    EPU index is correlated with a decrease in economic activity
  \end{itemize}
\end{itemize}
# Motivation and Methodology
## Motivation

A growing number of papers develop a country and region-specific EPU indexes

\vfill

\begin{itemize}
\setlength\itemsep{2em}
  \item
  Developed countries : Australia, Belgium, Canada, Denmark, France,
  Germany, Greece, Hong Kong, Ireland, Italy, Japan, Mexico, Singapore,
  South Korea, Spain, Sweden, The Netherlands, United Kingdom, United States
	\item
  Emerging markets :Brazil, Chile, China, Colombia, Croatia, India,
  Mexico, Russia, Pakistan
\end{itemize}

\vfill

We construct the Israeli Economic Policy Uncertainty index


## Lit Review

Literature suggests that fluctuations in uncertainty might have a significant impact on economic activity

\vfill

\begin{itemize}
  \item
  Employment and industrial investment
  \newline
  {\footnotesize	(Bloom 2009)}
  \item
  Monetary policy transmission and inflation expectations 
  \newline
  {\footnotesize (Aastveit, Natvik, and Sola 2017)}
  \item
  Bank credit growth 
  \newline
  {\footnotesize (Bordo, Duca, and Koch 2016)}
  \item
  Financial stability 
  \newline
  {\footnotesize (Phan et al. 2021)}
  \item
  De-anchorage of long-term in inflation expectations 
  \newline
  {\footnotesize (Dash, Rohit, and Devaguptapu 2020)}
\end{itemize}


## Methodology

  Identify relevant articles (using the same phrases as Baker, Bloom, and Davis
  (2016))

  \vfill

 \begin{itemize}
 \setlength\itemsep{2em}
 \item
 \textbf{Economic} - economy, economic
 \item
 \textbf{Policy} - prime minister, deficit, bank of israel, boi, knesset,
 parliament, regulation, deregulation, legislation, taxes, tax, taxation,
 taxed, policy,spending, budget.
 \item
 \textbf{Uncertainty} - uncertainty, uncertain, not certain.
 \end{itemize}

## Methodology

\begin{quote}
"... the \textcolor{blue}{economy} entered a continuing slowdown starting in the second half of 1987. this dumpened the public outlook on the future and caused increasing \textcolor{blue}{uncertainty}... 
 the governments savings dropped during the same years, due to the slowdown which cut \textcolor{blue}{tax} revenue from business and wage earners"
\end{quote}

\vspace{1em}

\begin{enumerate}
\setlength\itemsep{2em}
  \item
  Calculate the proportion of relevant articles in each time period 
  \begin{itemize}
    \item
    Monthly
    \item
    Quarterly
  \end{itemize}
  \item
  Scale to a mean of 100 
\end{enumerate}

## Data

\vspace{1em}

\begin{itemize}
\setlength\itemsep{2em}
  \item
  The Jerusalem Post 
  \newline
  (Long and consistent daily coverage of Israeli news)
  \item
  About 500,000 articles (about 50 a day)
  \item
  1988-2021 period
  \item
  Available from ProQuest
\end{itemize}

# Israeli EPU
## EPU seems to capture important domestic and international events

```{r plot_epu}

epu_month %>% 
  mutate(ma = slide_dbl(epu, mean,.before = 1 * 12)) %>% 
  pivot_longer(-date) %>% 
  mutate(name = factor(name, labels = c("EPU",
                                        "EPU moving average (1 year)"))) %>% 
  ggplot() +
  geom_line(aes(x = date, y = value, color = name)) +
  scale_color_manual(values = c("black", "blue")) + 
  geom_rect(data = filter(events_df,category == "crisis"),
            aes(xmin = start_date,xmax = end_date,ymin = -Inf,ymax = Inf),
            alpha = 0.5,fill = "lightblue") +
  geom_text(data = filter(events_df,category %in% c("political",
                                                    "international")),
            aes(x = start_date, y = c(450,550,450,550), label = event),
            color = "blue") +
  geom_segment(data = events_df %>% 
              filter(category %in% c("political",
                                     "international")) %>% 
              left_join(epu_month, by = c("start_date" = "date")),
              aes(x = start_date, y = c(430,530,430,530),
                  xend = start_date, yend = epu + 10),
              color = "blue", arrow = arrow(length = unit(4, "points"))) +
  xlab(NULL) + ylab(NULL) +
  ggtitle("Economic policy uncertainty index for Israel") + 
  labs(caption = glue("Shaded areas are periods of significant",
                      " economic slowdown in Israel")) + 
  theme(plot.caption = element_text(hjust = 0),
        legend.title = element_blank())
  

```

## EPU tends to move together with other measures of uncertainty


```{r benchmark_correlation}

full_period_plot = benchmark_df %>% 
  dplyr::select(-date,israel = epu) %>%
  rename_all(~if_else(. == "vix","Israeli VIX", str_to_title(.))) %>% 
  cor(use = "pairwise.complete.obs") %>% 
  ggcorrplot::ggcorrplot(type = "lower",
                         show.diag = FALSE,
                         lab = TRUE,
                         ggtheme = theme_light()) + 
  ggtitle("Full Period")

after_gfc_plot = benchmark_df %>% 
  filter(date >= as_date("2010-01-01")) %>% 
  dplyr::select(-date,israel = epu) %>%
  rename_all(~if_else(. == "vix","Israeli VIX", str_to_title(.))) %>% 
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot::ggcorrplot(type = "lower",
                         show.diag = FALSE,
                         lab = TRUE,
                         ggtheme = theme_light()) + 
  ggtitle("After GFC")

legend_row = cowplot::get_legend(
  full_period_plot + 
    theme(legend.box.margin = margin(0, 0, 0, 12)))

cowplot::plot_grid(full_period_plot + theme(legend.position = "none"),
                   after_gfc_plot + theme(legend.position = "none"),
                   legend_row,nrow = 1,scale = 1,axis = "l",
                   rel_widths = c(2,2,0.5))


```

## Israel EPU and Israeli VIX move together but also capture different aspects of uncertainty

```{r plot_israel_epu_vs_vix}


benchmark_df %>% 
  select(date,epu,vix) %>% 
  filter(complete.cases(.)) %>% 
  mutate(across(-date, scale)) %>% 
  pivot_longer(-date) %>% 
  mutate(name = factor(name, labels = c("Israel EPU", "Israel VIX"))) %>% 
  ggplot(aes(x = date,
             y = value)) + 
  geom_line(aes(color = name)) + 
  xlab(NULL) + ylab(NULL) + 
  theme(legend.title = element_blank())


```


## Topics EPU

\textbf{Economic} and \textbf{Policy} and \textbf{Uncertainty} dictionaries and :

\vfill

\begin{itemize}
\setlength\itemsep{2em}
  \item
  \textbf{Fiscal} - budget, deficit, government debt, tax, taxes, taxation,
  taxed, fiscal, military spending, defense spending, balanced budget,
  national debt.
  \item
  \textbf{Monetary} - boi, bank of israel, money supply, open market
  operations, quantitative easing, monetary policy, policy rate.
\end{itemize}

  
  
  


## High levels of Monetary and Fiscal uncertainty during "local" crises

```{r plot_epu_topics}

ggplot() +
  geom_line(data = epu_month_topics %>%
              filter(date >= as_date("2000-01-01")) %>% 
              mutate(name = factor(name,
                                   labels = c("Fiscal EPU", "Monetary EPU"))),
            aes(x = date, y = value, color = name)) +
  geom_rect(data = filter(events_df, category == "crisis"),
            aes(xmin = start_date,xmax = end_date,
                ymin = -Inf,ymax = Inf),
            alpha = 0.5, fill = "lightblue") +
  xlab(NULL) + ylab(NULL) +
  ggtitle("Economic policy uncertainty index for Israel") + 
  theme(legend.title = element_blank())

```


## Correlation between Monetary and Fiscal uncertainty drops after 2012

```{r plot_rolling_cor}

epu_month_topics %>%
  pivot_wider(id_cols = date) %>%
  mutate(roll_cor = slide2_dbl(
    .x = monetary_epu,
    .y = fiscal_epu,
    .f = ~ cor(.x, .y),
    .before = 10*12,
    .complete = TRUE
  )) %>% 
  filter(complete.cases(.)) %>% 
  ggplot(aes(x = date, y = roll_cor)) + 
  geom_line() + 
  xlab(NULL) + ylab(NULL) + 
  ggtitle(glue("10 year Rolling Correlations",
               " Between Monetary and Fiscal EPU indices"))

```

```{r plot_rolling_cor_robust, eval=FALSE}

epu_month_topics %>%
  pivot_wider(id_cols = date) %>%
  mutate(roll_cor_2 = slide2_dbl(
    .x = monetary_epu,
    .y = fiscal_epu,
    .f = ~ cor(.x, .y),
    .before = 2*12,
    .complete = TRUE
  )) %>% 
  mutate(roll_cor_10 = slide2_dbl(
    .x = monetary_epu,
    .y = fiscal_epu,
    .f = ~ cor(.x, .y),
    .before = 10*12,
    .complete = TRUE
  )) %>%
  # filter(complete.cases(.)) %>% 
  select(date, contains("roll_cor")) %>% 
  pivot_longer(-date) %>% 
  ggplot(aes(x = date, y = value, color = name)) + 
  geom_line() + 
  xlab(NULL) + ylab(NULL)

 ```


# Israeli EPU and economic activity
## The effect of EPU on economic activity - SVAR model

\begin{equation*}
  Z_{t} = B Z_{t - 1} + u_{t}
\end{equation*}

where:

Z - vector of (ordered by Cholesky decomposition) economic policy uncertainty,
the employment rate, the nominal effective exchange rate, the nominal interest
rate and the GDP

\vfill

B - coefficient matrix at 5 lags (according to AIC)

\vfill

$u_{t}$ - set of errors with mean zero, zero autocorrelation, and variance covariance matrix $\Sigma = E[u_{t}u_{t}']$



## GDP responds negatively to uncertainty shocks

```{r plot_log_levels}


map_dfr(c("irf", "Lower", "Upper"), function(temp_name) {
  return(
    irf_df %>%
      pluck(temp_name, 1) %>%
      as_tibble() %>%
      rownames_to_column(var = "lag") %>%
      pivot_longer(-lag) %>%
      mutate(category = temp_name)
  )
  
  
}) %>%
  pivot_wider(names_from = category) %>%
  rename_all(tolower) %>%
  filter(!name == "epu_log") %>%
  mutate(name = factor(name,
    levels = c("gdp_il_sa_log", "emp_il_sa_log", "i_il_log",
               "fx_eff_log","epu_log"),
    labels = c("GDP", "Employment", "Interest rate", "Exchange rate",
               "EPU"))) %>% 
  ggplot() +
  geom_line(aes(x = lag, y = irf, group = 1)) +
  geom_ribbon(aes(x = lag, ymax = upper, ymin = lower, group = 1),
              fill = "lightblue",
              alpha = 0.2) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  scale_y_continuous(labels = scales::percent_format()) + 
  xlab(NULL) + ylab(NULL) + 
  ggtitle("Effect of orthogonal EPU shock (log levels)") + 
  facet_wrap( ~ name, scales = "free")




```

## Robustness 

```{r plot_log_delta}

map_dfr(c("irf", "Lower", "Upper"), function(temp_name) {
  return(
    var_irf_delta %>%
      pluck(temp_name, 1) %>%
      as_tibble() %>%
      rownames_to_column(var = "lag") %>%
      pivot_longer(-lag) %>%
      mutate(category = temp_name)
  )
  
  
}) %>%
  pivot_wider(names_from = category) %>%
  rename_all(tolower) %>%
  filter(!name == "epu") %>%
  mutate(name = factor(name,
    levels = c("gdp_il_sa", "emp_il_sa", "i_il", "fx_eff", "epu"),
    labels = c("GDP", "Employment", "Interest rate", "Exchange rate",
               "EPU"))) %>% 
  ggplot() +
  geom_line(aes(x = lag, y = irf, group = 1)) +
  geom_ribbon(aes(x = lag, ymax = upper, ymin = lower, group = 1),
              fill = "lightblue",
              alpha = 0.2) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  scale_y_continuous(labels = scales::percent_format()) + 
  xlab(NULL) + ylab(NULL) + 
  ggtitle("Effect of orthogonal EPU shock (log changes)") + 
  facet_wrap( ~ name, scales = "free")




```


# Conclusion
## Conclusion

\vfill

\begin{itemize}
\setlength\itemsep{2em}
  \item
  We construct an Economic Policy Uncertainty Index for Israel (and Monetary
  and Fiscal categories).
  \item
  The Israeli EPU index reacts to major local and international events
  and correlates with other uncertainty measures.
  \item
  The results suggest that Israel EPU may be used for forecasting and monitoring
  the Israeli economy.
\end{itemize}

# Appendix

```{r import_foreign_vix}

file_path = paste0(Sys.getenv("USERPROFILE"),
                   "\\OneDrive - Bank Of Israel",
                   "\\Data\\EPU\\vix\\VIX.xlsx")

foreign_vix = map_dfr(c("euro","us"), function(temp_ind){
  
  temp_df = read_xlsx(file_path, sheet = temp_ind)
  
  temp_df = temp_df %>% 
    mutate(date = as_date(date)) %>% 
    rename(value = 2) %>% 
    mutate(country = temp_ind)
  
  
})


```

## US VIX

```{r plot_us_vix}

epus_df %>% 
  select(date, usa) %>% 
  inner_join(foreign_vix %>% 
               filter(country == "us") %>% 
               select(-country)) %>% 
  mutate(across(-date, scale)) %>% 
  pivot_longer(-date) %>% 
  mutate(name = factor(name, labels = c("US EPU", "US VIX"))) %>% 
  ggplot(aes(x = date,
             y = value)) + 
  geom_line(aes(color = name)) + 
  xlab(NULL) + ylab(NULL) + 
  theme(legend.title = element_blank())

```

## Euro VIX

```{r plot_euro_vix}

epus_df %>% 
  select(date, euro) %>% 
  inner_join(foreign_vix %>% 
               filter(country == "euro") %>% 
               select(-country)) %>% 
  mutate(across(-date, scale)) %>% 
  pivot_longer(-date) %>% 
  mutate(name = factor(name, labels = c("Euro EPU", "Euro VIX"))) %>% 
  ggplot(aes(x = date,
             y = value)) + 
  geom_line(aes(color = name)) + 
  xlab(NULL) + ylab(NULL) + 
  theme(legend.title = element_blank())

```

## Euro vix calculations

 We draw on two newspapers per country for the European indexes: Le Monde and Le Figaro
 for **France**, Handelsblatt and Frankfurter Allgemeine Zeitung for **Germany**,
 Corriere Della Sera and La Stampa for **Italy**, El Mundo and El Pais for **Spain**,
 and The Times of London and Financial Times for the **United Kingdom**.
 
 The index holds stocks from nine eurozone countries: Belgium, Finland, France, Germany,
 Ireland, Italy, Luxembourg, the Netherlands, and Spain
