---
title: "Measuring Economic Policy Uncertainty for the Israeli Economy"
author: "Yaakov Chen-Zion, Michael Gurkov and Nimrod Segev"
institute: "Bank of Israel \\newline Research Department \\newline \\newline Any views expressed in the Discussion Paper are those of the authors and do not necessarily reflect those of the Bank of Israel"
date: "16 June, 2021"
classoption: t
output:
  beamer_presentation:
    includes:
      in_header: !expr here::here('PresentationPreamble.tex')
    latex_engine: xelatex
    slide_level: 2
# bibliography: "GaR.bib"
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, echo = FALSE, message = FALSE,
  warning = FALSE, comment = "#>"

)
```


```{r load_libraries}

library(lubridate)

library(glue)

library(xts)

library(slider)

library(readxl)

library(corrr)

library(vars)

library(aTSA)

library(tidyverse)

```


```{r set_params}

events_df = tribble(~event, ~start_date, ~end_date,
                    "Intifada","2001-01-01","2003-01-01",
                    "GFC","2008-01-01","2009-01-01",
                    "Coona","2020-01-01","2021-01-01") %>% 
  mutate(across(ends_with("date"),as_date))


```


```{r import_data_epu}

epu_month = read_csv(glue(Sys.getenv("USERPROFILE"),
                       "\\OneDrive - Bank Of Israel\\Data\\",
                       "EPU\\TDM-JP\\temp_epu_month.csv")) %>% 
  dplyr::select(date = time_period, epu = epu_scaled)


epu_month_topics = read_csv(glue(Sys.getenv("USERPROFILE"),
                       "\\OneDrive - Bank Of Israel\\Data\\",
                       "EPU\\TDM-JP\\temp_epu_month_topics.csv")) %>% 
  dplyr::select(date = time_period, monetary_epu = monetary_epu_scaled,
         fiscal_epu = fiscal_epu_scaled) %>%
  mutate(date = as_date(paste0(date,"-01"))) %>% 
  pivot_longer(-date)



```

```{r import_epus}

url_table = tribble(
  ~ name, ~ file_name,~ var_name,
  "usa","US_Policy_Uncertainty_Data","News_Based_Policy_Uncert_Index",
  "global","Global_Policy_Uncertainty_Data", "GEPU_current"
)

epus_df = pmap(url_table,function(name, file_name, var_name){
  
  temp_df = read_xlsx(glue(Sys.getenv("USERPROFILE"),
                           "\\OneDrive - Bank Of Israel\\Data\\EPU",
                           "\\epus\\",file_name, ".xlsx"))
  
  temp_df = temp_df %>% 
    mutate(Month = str_pad(Month,side = "left",width = 2,pad = "0")) %>% 
    mutate(date = as_date(paste(Year,Month,"01", sep = "-"))) %>% 
    dplyr::select(date, !!sym(name) := all_of(var_name))
      
  
  return(temp_df)

  
}) %>% 
  reduce(inner_join, "date")

rm(url_table)

```

```{r import_var_data}

vars_list = c("epu", "fx_us", "ta_125", "i_il", "emp_il_sa",
              "gdp_il_sa","fx_eff")

var_df = read_excel(glue(Sys.getenv("USERPROFILE"),
                         "\\OneDrive - Bank Of Israel",
                         "\\Data\\EPU\\EPUI",
                         "\\EPUI_Data_activity_ALL_q.xlsx"),
                    sheet = "DATA")

var_df = var_df %>%
  mutate(date = as.yearqtr(paste0(year,"Q",quarter))) %>% 
  select(date,any_of(vars_list),starts_with("fx")) %>%
  mutate(fx_eff = if_else(year(date) > 1999,
                          fx_ner_ef,
                          0.75 * fx_us + 0.07 * fx_japan + 0.18 * fx_uk)) %>%
  mutate(across(vars_list,log)) %>%
  mutate(across(vars_list, lag,.names = "l_{.col}")) %>% 
  filter(date >= as.yearqtr("1999Q2") & year(date) <= 2019)

rm(vars_list)

```



## Outline

\tableofcontents[sectionstyle=show/show,subsectionstyle=hide/hide]

## Motivation

\begin {itemize}
\itemsep3em
	\item
	Literature suggests that fluctuations in uncertainty
	might have a significant impact on real economic activity
	(\textcolor{red}{add citation}) (Bloom 2009; Caggiano,
	Castelnuovo, and Groshenny 2014; Caldara et al. 2016)
	\item
	
	\item
	We construct the Israeli Economic Policy Uncertainty index


\end{itemize}



## Methodology

Data : about 500,000 articles for the 1988-2021 period from the Jerusalem Post
  long and consistent daily coverage of Israeli news
  exact same phrases used by Baker, Bloom, and Davis (2016)

 * Economic - economy, economic
 * Policy - prime minister, deficit, Bank of Israel, BoI, knesset,
    parliament, regulation, deregulation, legislation, taxes, tax, taxation,
    taxed, policy,spending, budget.
 * Uncertainty - uncertainty, uncertain, not certain.

## Methodology

\begin{enumerate}
  \item
  Calculate the proportion of "relevant" articles in each time period (month/quarter)
  \item
  Standardize to unit standard deviation and normalize to a mean of 100 
\end{enumerate}

##

```{r plot_epu}

ggplot() + 
  geom_line(data = epu_month, aes(x = date, y = epu)) + 
  geom_rect(data = events_df, aes(xmin = start_date, xmax = end_date,
                                  ymin = -Inf, ymax = Inf),
            alpha = 0.5, fill = "lightblue") +
  xlab(NULL) + ylab(NULL) + 
  ggtitle("Economic policy uncertainty index for Israel")

```

##

```{r benchmark_correlation}

full_period_plot = epu_month %>%
  inner_join(epus_df, "date") %>%
  dplyr::select(-date,israel = epu) %>%
  filter(complete.cases(.)) %>%
  cor() %>%
  ggcorrplot::ggcorrplot(type = "lower",
                         show.diag = FALSE,
                         lab = TRUE,
                         ggtheme = theme_light()) + 
  ggtitle("Full Period")

after_gfc_plot = epu_month %>%
  inner_join(epus_df, "date") %>%
  filter(date >= as_date("2009-01-01")) %>% 
  dplyr::select(-date,israel = epu) %>%
  filter(complete.cases(.)) %>%
  cor() %>%
  ggcorrplot::ggcorrplot(type = "lower",
                         show.diag = FALSE,
                         lab = TRUE,
                         ggtheme = theme_light()) + 
  ggtitle("After GFC")

legend_row = cowplot::get_legend(full_period_plot + 
                                   theme(legend.box.margin = margin(0, 0, 0, 12)))

cowplot::plot_grid(full_period_plot + theme(legend.position = "none"),
                   after_gfc_plot + theme(legend.position = "none"),
                   legend_row,nrow = 1,scale = 1,axis = "l",
                   rel_widths = c(2,2,0.5))


```

## Topics EPU

  * Fiscal - budget, deficit, government debt, tax, taxes, taxation,
  taxed, fiscal, military spending, defense spending, balanced budget,
  national debt.
  
  * Monetary - boi, bank of israel, money supply, open market
  operations, quantitive easing, monetary policy, policy rate.


```{r plot_epu_topics}

ggplot() +
  geom_line(data = epu_month_topics %>%
              mutate(name = factor(name,
                                   labels = c("Fiscal EPU", "Monetary EPU"))),
            aes(x = date, y = value, color = name)) +
  geom_rect(data = events_df,aes(xmin = start_date,xmax = end_date,
                                 ymin = -Inf,ymax = Inf),
            alpha = 0.5, fill = "lightblue") +
  xlab(NULL) + ylab(NULL) +
  ggtitle("Economic policy uncertainty index for Israel") + 
  theme(legend.title = element_blank())

```


##

```{r plot_rolling_cor}

epu_month_topics %>%
  pivot_wider(id_cols = date) %>%
  mutate(roll_cor = slide2_dbl(
    .x = monetary_epu,
    .y = fiscal_epu,
    .f = ~ cor(.x, .y),
    .before = 10*12,
    .complete = TRUE
  )) %>% 
  filter(complete.cases(.)) %>% 
  ggplot(aes(x = date, y = roll_cor)) + 
  geom_line() + 
  xlab(NULL) + ylab(NULL) + 
  ggtitle(glue("10 year Rolling Correlations",
               " Between Monetary and Fiscal EPU indices"))

```

## VAR

\begin{equation}
  Z_{t} = B Z_{t - 1} + u_{t}
\end{equation}

where:

Z - vector of economic policy uncertainty, the employment rate,
the nominal effective exchange rate, the nominal interest rate and
the GDP

B - coefficient matrix

$u_{t}$ - set of errors with mean zero, zero autocorrelation, and variance covariance matrix $\Sigma = E[u_{t}u_{t}']$


## VAR estimation results

```{r }

var_irf =  var_df  %>%
  dplyr::select(epu, fx_eff, i_il, emp_il_sa, gdp_il_sa)  %>%
  ts(start = c(1995, 2), frequency = 4)  %>%
  VAR(p = 8, type = "both")  %>%
  irf(n.ahead = 6,runs = 1000,ci = 0.90,impulse = "epu")


map_dfr(c("irf", "Lower", "Upper"), function(temp_name) {
  return(
    var_irf %>%
      pluck(temp_name, 1) %>%
      as_tibble() %>%
      rownames_to_column(var = "lag") %>%
      pivot_longer(-lag) %>%
      mutate(category = temp_name)
  )
  
  
}) %>%
  pivot_wider(names_from = category) %>%
  rename_all(tolower) %>%
  filter(!name == "epu") %>%
  # mutate(lag = as.numeric(lag)) %>%
  mutate(name = factor(name,
    levels = c("gdp_il_sa", "emp_il_sa", "i_il", "fx_eff"),
    labels = c("GDP", "Employment", "Interest rate", "Exchange rate"))) %>% 
  ggplot() +
  geom_line(aes(x = lag, y = irf, group = 1)) +
  geom_ribbon(aes(x = lag, ymax = upper, ymin = lower, group = 1),
              fill = "lightblue",
              alpha = 0.5) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  xlab(NULL) + ylab(NULL) + ggtitle("Effect of orthogonal EPU shock") + 
  facet_wrap( ~ name, scales = "free")




```


```{r, eval=FALSE}

unlist_irf_list = function(temp_list){
  
  var_names = c("irf","Lower","Upper")
  
  temp_df = map_dfc(var_names, function(temp_name){
    
    return(temp_list %>% 
      pluck(temp_name) %>% 
      unlist(use.names = FALSE))
 
  }) %>% 
    set_names(var_names) %>% 
    rename_all(tolower)
  
  
  
  
}

```


```{r plot_var, eval=FALSE}

temp_df %>% 
  rownames_to_column(var = "lags") %>% 
  ggplot(aes(x = lags, y = irf, group = 1)) + 
  geom_line() + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  geom_ribbon(aes(x = lags, ymax = upper, ymin = lower),
              fill = "lightblue", alpha = 0.3)

```

