---
title: "Measuring Economic Policy Uncertainty for the Israeli Economy"
author: "Yaakov Chen-Zion, Michael Gurkov and Nimrod Segev"
institute: "Bank of Israel \\newline Research Department \\newline \\newline Any views expressed in the Discussion Paper are those of the authors and do not necessarily reflect those of the Bank of Israel"
date: "13 December, 2021"
classoption: t
output:
  beamer_presentation:
    includes:
      in_header: !expr here::here('PresentationPreamble.tex')
    latex_engine: xelatex
    slide_level: 2
# bibliography: "GaR.bib"
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, echo = FALSE, message = FALSE,
  warning = FALSE, comment = "#>"

)
```


```{r load_libraries}

library(lubridate)

library(glue)

library(xts)

library(slider)

library(readxl)

library(corrr)

library(vars)

library(aTSA)

library(tidyverse)

```


```{r set_params}

events_df = tribble(~event, ~category, ~start_date, ~end_date,
                    "Intifada","crisis","2001-01-01","2003-01-01",
                    "GFC","crisis","2008-01-01","2009-01-01",
                    "Corona","crisis","2020-01-01","2021-01-01",
                    "Ariel Sharon","political","2006-01-01","2006-02-01",
                    "Social Justice","political","2011-06-01","2011-12-01",
                    "Brexit","international","2016-06-01","2016-07-01",
                    "Trump","international","2016-11-01","2016-12-01") %>% 
  mutate(across(ends_with("date"),as_date))


```


```{r import_data_epu}

epu_month = read_csv(glue(Sys.getenv("USERPROFILE"),
                       "\\OneDrive - Bank Of Israel\\Data\\",
                       "EPU\\TDM-JP\\temp_epu_month.csv")) %>% 
  dplyr::select(date = time_period, epu = epu_scaled)


epu_month_topics = read_csv(glue(Sys.getenv("USERPROFILE"),
                       "\\OneDrive - Bank Of Israel\\Data\\",
                       "EPU\\TDM-JP\\temp_epu_month_topics.csv")) %>% 
  dplyr::select(date = time_period, monetary_epu = monetary_epu_scaled,
         fiscal_epu = fiscal_epu_scaled) %>%
  mutate(date = as_date(paste0(date,"-01"))) %>% 
  pivot_longer(-date)



```

```{r import_epus}

url_table = tribble(
  ~ name, ~ file_name,~ var_name,
  "usa","US_Policy_Uncertainty_Data","News_Based_Policy_Uncert_Index",
  "global","Global_Policy_Uncertainty_Data", "GEPU_current",
  "euro","Europe_Policy_Uncertainty_Data","European_News_Index"
)

epus_df = pmap(url_table,function(name, file_name, var_name){
  
  temp_df = read_xlsx(glue(Sys.getenv("USERPROFILE"),
                           "\\OneDrive - Bank Of Israel\\Data\\EPU",
                           "\\epus\\",file_name, ".xlsx"))
  
  temp_df = temp_df %>% 
    mutate(Month = str_pad(Month,side = "left",width = 2,pad = "0")) %>% 
    mutate(date = as_date(paste(Year,Month,"01", sep = "-"))) %>% 
    dplyr::select(date, !!sym(name) := all_of(var_name))
      
  
  return(temp_df)

  
}) %>% 
  reduce(full_join, "date")

rm(url_table)

```

```{r import_vix}

vix = read_csv(glue(Sys.getenv("USERPROFILE"),
                    "\\OneDrive - Bank Of Israel\\Data\\VIX.csv")) %>% 
  rename_all(tolower) %>% 
  select(date, vix = close)

vix = vix %>% 
  # group_by(date = as.yearmon(date)) %>% 
  # summarise(vix = mean(vix, na.rm = TRUE), .groups = "drop") %>% 
  # mutate(date = paste("01", str_pad(month(date),2,"left",0),
  #                          year(date),sep = "-")) %>% 
  # mutate(date = dmy(date)) %>% 
  filter(date >= as_date("2006-06-01"))


```



```{r import_var_data}

vars_list = c("epu", "i_il", "emp_il_sa",
              "gdp_il_sa","fx_eff")

var_df = read_excel(glue(Sys.getenv("USERPROFILE"),
                         "\\OneDrive - Bank Of Israel",
                         "\\Data\\EPU\\EPUI",
                         "\\EPUI_Data_activity_ALL_q.xlsx"),
                    sheet = "DATA")

var_df = var_df %>%
  mutate(date = as.yearqtr(paste0(year,"Q",quarter))) %>% 
  select(date,any_of(vars_list),starts_with("fx")) %>%
  mutate(fx_eff = if_else(year(date) > 1999,
                          fx_ner_ef,
                          0.75 * log(fx_us) + 0.07 * fx_japan + 0.18 * fx_uk)) %>%
  mutate(across(any_of(vars_list),log,.names = "{.col}_log")) %>%
  mutate(across(ends_with("_log"), ~ . - lag(.),.names = "{.col}_diff")) %>% 
  filter(date >= as.yearqtr("1995Q2") & year(date) <= 2019)


```

```{r run_var, cache=TRUE}

irf_df =  var_df  %>%
  dplyr::select(epu_log, fx_eff_log, i_il_log, emp_il_sa_log, gdp_il_sa_log)  %>%
  ts(start = c(1995, 2), frequency = 4)  %>%
  VAR(p = 8, ic = "AIC", type = "both", lag.max = 8)  %>%
  irf(n.ahead = 6,runs = 1000,ci = 0.90,impulse = "epu_log")

var_irf_delta =  var_df  %>%
  dplyr::select(paste0(vars_list,"_log_diff")) %>%
  dplyr::rename_all(~str_remove(.,"_log_diff")) %>% 
  filter(complete.cases(.)) %>% 
  ts(start = c(1995, 2), frequency = 4)  %>%
  VAR(p = 8, ic = "AIC", type = "both", lag.max = 8)  %>%
  irf(n.ahead = 6,cumulative = TRUE,runs = 1000,ci = 0.90,impulse = "epu")



```


## Outline

\tableofcontents[sectionstyle=show/show,subsectionstyle=hide/hide]

## Motivation

Literature suggests that fluctuations in uncertainty might have a significant impact on economic activity

\vfill

\begin{itemize}
  \item
  Employment and industrial investment
  \newline
  (Bloom 2009)
  \item
  Monetary policy transmission and inflation expectations 
  \newline
  (Aastveit, Natvik, and Sola 2017)
  \item
  Bank credit growth 
  \newline
  (Bordo, Duca, and Koch 2016)
  \item
  Financial stability 
  \newline
  (Phan et al. 2021)
  \item
  De-anchorage of long-term in inflation expectations 
  \newline
  (Dash, Rohit, and Devaguptapu 2020)
\end{itemize}

## Motivation

A growing number of papers develop a country and region-specific EPU indexes

\begin{itemize}
	\item
	USA
	\item
	UK
	\item
	Canada
	\item
	Germany
	\item
	France
	\item
	Sweden
	\item
	South Korea
	\item
	Singapore
\end{itemize}

We construct the Israeli Economic Policy Uncertainty index


## Methodology

Data : about 500,000 articles for the 1988-2021 period from the Jerusalem Post.
  Long and consistent daily coverage of Israeli news
  Exact same phrases used by Baker, Bloom, and Davis (2016)

 \begin{itemize}
 \setlength\itemsep{2em}
 \item
 Economic - economy, economic
 \item
 Policy - prime minister, deficit, Bank of Israel, BoI, knesset,
 parliament, regulation, deregulation, legislation, taxes, tax, taxation,
 taxed, policy,spending, budget.
 \item
 Uncertainty - uncertainty, uncertain, not certain.
 \end{itemize}

## Methodology

Economic Policy Uncertainty Index calculation:

\vfill

\begin{enumerate}
\setlength\itemsep{2em}
  \item
  Calculate the proportion of "relevant" articles in each time period 
  \begin{itemize}
    \item
    Monthly
    \item
    Quarterly
  \end{itemize}
  \item
  Standardize to unit standard deviation
  \item
  Scale to a mean of 100 
\end{enumerate}

## EPU seems to capture important domestic and international events

```{r plot_epu}


ggplot() +
  geom_line(data = epu_month, aes(x = date, y = epu)) +
  geom_rect(data = filter(events_df,category == "crisis"),
            aes(xmin = start_date,xmax = end_date,ymin = -Inf,ymax = Inf),
            alpha = 0.5,fill = "lightblue") +
  geom_text(data = filter(events_df,category == "political"),
            aes(x = start_date, y = c(450,550), label = event),
            color = "blue") +
  geom_segment(data = events_df %>% 
              filter(category == "political") %>% 
              left_join(epu_month, by = c("start_date" = "date")),
              aes(x = start_date, y = c(430,530),
                  xend = start_date, yend = epu + 10),
              color = "blue", arrow = arrow(length = unit(4, "points"))) +
  xlab(NULL) + ylab(NULL) +
  ggtitle("Economic policy uncertainty index for Israel") + 
  labs(caption = glue("Shaded areas are periods of significant",
                      " economic slowdown in Israel")) + 
  theme(plot.caption = element_text(hjust = 0))
  

```

## EPU tends to move together with other measures of uncertainty


```{r benchmark_correlation}

benchmark_df = epu_month %>%
  full_join(epus_df, "date") %>%
  full_join(vix, "date")

full_period_plot = benchmark_df %>% 
  dplyr::select(-date,israel = epu) %>%
  rename_all(~if_else(. == "vix","VIX", str_to_title(.))) %>% 
  cor(use = "pairwise.complete.obs") %>% 
  ggcorrplot::ggcorrplot(type = "lower",
                         show.diag = FALSE,
                         lab = TRUE,
                         ggtheme = theme_light()) + 
  ggtitle("Full Period")

after_gfc_plot = benchmark_df %>% 
  filter(date >= as_date("2010-01-01")) %>% 
  dplyr::select(-date,israel = epu) %>%
  rename_all(~if_else(. == "vix","VIX", str_to_title(.))) %>% 
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot::ggcorrplot(type = "lower",
                         show.diag = FALSE,
                         lab = TRUE,
                         ggtheme = theme_light()) + 
  ggtitle("After GFC")

legend_row = cowplot::get_legend(
  full_period_plot + 
    theme(legend.box.margin = margin(0, 0, 0, 12)))

cowplot::plot_grid(full_period_plot + theme(legend.position = "none"),
                   after_gfc_plot + theme(legend.position = "none"),
                   legend_row,nrow = 1,scale = 1,axis = "l",
                   rel_widths = c(2,2,0.5))


```



## Topics EPU

  * Fiscal - budget, deficit, government debt, tax, taxes, taxation,
  taxed, fiscal, military spending, defense spending, balanced budget,
  national debt.
  
  * Monetary - boi, bank of israel, money supply, open market
  operations, quantitive easing, monetary policy, policy rate.


## High levels of Monetary and Fiscal uncertainty during "local" crises

```{r plot_epu_topics}

ggplot() +
  geom_line(data = epu_month_topics %>%
              mutate(name = factor(name,
                                   labels = c("Fiscal EPU", "Monetary EPU"))),
            aes(x = date, y = value, color = name)) +
  geom_rect(data = events_df,aes(xmin = start_date,xmax = end_date,
                                 ymin = -Inf,ymax = Inf),
            alpha = 0.5, fill = "lightblue") +
  xlab(NULL) + ylab(NULL) +
  ggtitle("Economic policy uncertainty index for Israel") + 
  theme(legend.title = element_blank())

```


## Corellation between Monetary and Fiscal uncertainty drops after 2012

```{r plot_rolling_cor}

epu_month_topics %>%
  pivot_wider(id_cols = date) %>%
  mutate(roll_cor = slide2_dbl(
    .x = monetary_epu,
    .y = fiscal_epu,
    .f = ~ cor(.x, .y),
    .before = 10*12,
    .complete = TRUE
  )) %>% 
  filter(complete.cases(.)) %>% 
  ggplot(aes(x = date, y = roll_cor)) + 
  geom_line() + 
  xlab(NULL) + ylab(NULL) + 
  ggtitle(glue("10 year Rolling Correlations",
               " Between Monetary and Fiscal EPU indices"))

```

## VAR

\begin{equation}
  Z_{t} = B Z_{t - 1} + u_{t}
\end{equation}

where:

Z - vector of economic policy uncertainty, the employment rate,
the nominal effective exchange rate, the nominal interest rate and
the GDP

B - coefficient matrix

$u_{t}$ - set of errors with mean zero, zero autocorrelation, and variance covariance matrix $\Sigma = E[u_{t}u_{t}']$


## VAR estimation results

```{r plot_log_levels}


map_dfr(c("irf", "Lower", "Upper"), function(temp_name) {
  return(
    irf_df %>%
      pluck(temp_name, 1) %>%
      as_tibble() %>%
      rownames_to_column(var = "lag") %>%
      pivot_longer(-lag) %>%
      mutate(category = temp_name)
  )
  
  
}) %>%
  pivot_wider(names_from = category) %>%
  rename_all(tolower) %>%
  filter(!name == "epu_log") %>%
  mutate(name = factor(name,
    levels = c("gdp_il_sa_log", "emp_il_sa_log", "i_il_log", "fx_eff_log"),
    labels = c("GDP", "Employment", "Interest rate", "Exchange rate"))) %>% 
  ggplot() +
  geom_line(aes(x = lag, y = irf, group = 1)) +
  geom_ribbon(aes(x = lag, ymax = upper, ymin = lower, group = 1),
              fill = "lightblue",
              alpha = 0.2) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  xlab(NULL) + ylab(NULL) + 
  ggtitle("Effect of orthogonal EPU shock (log levels)") + 
  facet_wrap( ~ name, scales = "free")




```

## VAR estimation results

```{r plot_log_delta}

map_dfr(c("irf", "Lower", "Upper"), function(temp_name) {
  return(
    var_irf_delta %>%
      pluck(temp_name, 1) %>%
      as_tibble() %>%
      rownames_to_column(var = "lag") %>%
      pivot_longer(-lag) %>%
      mutate(category = temp_name)
  )
  
  
}) %>%
  pivot_wider(names_from = category) %>%
  rename_all(tolower) %>%
  filter(!name == "epu") %>%
  mutate(name = factor(name,
    levels = c("gdp_il_sa", "emp_il_sa", "i_il", "fx_eff"),
    labels = c("GDP", "Employment", "Interest rate", "Exchange rate"))) %>% 
  ggplot() +
  geom_line(aes(x = lag, y = irf, group = 1)) +
  geom_ribbon(aes(x = lag, ymax = upper, ymin = lower, group = 1),
              fill = "lightblue",
              alpha = 0.2) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  xlab(NULL) + ylab(NULL) + 
  ggtitle("Effect of orthogonal EPU shock (log levels)") + 
  facet_wrap( ~ name, scales = "free")




```

